---
title: "Plants & Profits: Portfolio 1, Part 1.4"
output: html_notebook
---

Portfolio 1: Part 1.4
By Raefah Wahid

Part 1: Maps

```{r}
library("BIEN")
library(ape) #Package for working with phylogenies in R
library(maps) #Useful for making quick maps of occurrences
library(sp) # A package for spatial data
library(dplyr)
library(lattice)
```

The species for my book, The Signature of All Things, is Dicranum scoparium. Here is the map of range and occurrence for that species of moss:

```{r}
moss_range <- BIEN_ranges_load_species("Dicranum scoparium")
moss_occurrence <- BIEN_occurrence_species(species = "Dicranum scoparium")
map('world', fill=T, col="grey", bg="light blue", xlim=c(-180,-20), ylim=c(-60, 80))
plot(moss_range, col="green", add=T)
points(cbind(moss_occurrence$longitude, moss_occurrence$latitude), col="blue", pch=20, cex=1)
```

For the second species, I decided to go with the Cinchona calisaya, which is also featured in Elizabeth Gilbert's The Signature of All Things and is the focus of my essay.

```{r}
cinchona_range <- BIEN_ranges_load_species("Cinchona calisaya")
cinchona_occurrence <- BIEN_occurrence_species(species = "Cinchona calisaya")
map('world', fill=T, col="grey", bg="light blue", xlim=c(-180,-20), ylim=c(-60, 80))
plot(cinchona_range, col="green", add=T)
points(cbind(cinchona_occurrence$longitude, cinchona_occurrence$latitude), col="blue", pch=20, cex=1)
```

The two plants are in totally different areas! That's so cool, and makes sense, since the cinchona requires a hot climate (like in South America) while moss is prevalent in North America.

Part 2: Trait Data

From the exercises, I chose to do the following question: Do you think the difference in whole plant height is statistically significant between Artocarpus hirsutus and Artocarpus integer? But there was no whole plant height data for Artocarpus integer (at least, it wasn't coming up for me), so I just compared Artocarpus hirsutus with Artocarpus altilis.

The following snippet of code creates a dataframe by the name of Artocarpus_hirsutus, which has information about the whole plant height of the species Artocarpus hirsutus.   

```{r}
Artocarpus_hirsutus <- BIEN_trait_traitbyspecies(trait = "whole plant height", species = "Artocarpus hirsutus")
str(Artocarpus_hirsutus)
Artocarpus_hirsutus
```

The following snippet of code creates a data frame by the name of Artocarpus_altilis, which has information about the whole plant height of the species Artocarpus altilis.

```{r}
Artocarpus_altilis <- BIEN_trait_traitbyspecies(trait = "whole plant height", species = "Artocarpus altilis")
str(Artocarpus_altilis)
Artocarpus_altilis
```

Now, I'm going to compute the mean whole plant height of both species.

```{r}
hirsutus = Artocarpus_hirsutus[c(3)] # looking at column 3 of data, which is WPH
hirsutus = as.data.frame(hirsutus)
hirsutus_height = as.numeric(hirsutus$trait_value)
mean(hirsutus_height, data=hirsutus)

altilis = Artocarpus_altilis[c(3)] # looking at column 3 of data, which is WPH
altilis = as.data.frame(altilis)
altilis_height = as.numeric(altilis$trait_value)
mean(altilis_height, data=altilis)
```

The mean whole plant height of Artocarpus hirsutus is approximately 9.6 m. The mean whole plant height of Artocarpus altilis is approximatiely 11.5 m. This doesn't seem like a very signifianct difference in whole plant height, but we can examine this difference further using a box-plot.

```{r}
# making boxplots
boxplot(hirsutus_height)
boxplot(altilis_height)
```

Let's make a t-test to further analyze the average whole plant height of the two species.

```{r}
# computing t-test
t.test(hirsutus_height, altilis_height)
```

Now that we've done our t-test, we can see that the p-value = 0.1302, which is much greater than 0.05. Thus, the t-test shows us that the difference in whole plant height between these two species (Artocarpus hirsutus and Artocarpus atilis) isn't very significant at all.

Let's try this again for two other species. This time, let's use avocados!  

The following snippet of code creates a dataframe by the name of Persea_americana, which has information about the whole plant height of the species Persea americana.   

```{r}
Persea_americana <- BIEN_trait_traitbyspecies(trait = "whole plant height", species = "Persea americana")
str(Persea_americana)
Persea_americana
```

The following snippet of code creates a data frame by the name of Persea_borbonia, which has information about the whole plant height of the species Persea borbonia.

```{r}
Persea_borbonia <- BIEN_trait_traitbyspecies(trait = "whole plant height", species = "Persea borbonia")
str(Persea_borbonia)
Persea_borbonia
```

Now, I'm going to compute the mean whole plant height of both species.

```{r}
americana = Persea_americana[c(3)] # looking at column 3 of data, which is WPH
americana = as.data.frame(americana)
americana_height = as.numeric(americana$trait_value)
mean(americana_height, data=americana)

borbonia = Persea_borbonia[c(3)] # looking at column 3 of data, which is WPH
borbonia = as.data.frame(borbonia)
borbonia_height = as.numeric(borbonia$trait_value)
mean(borbonia_height, data=borbonia)
```

The mean whole plant height of Persea americana is approximately 11.6 m. The mean whole plant height of Persea borbonia is approximatiely 8.2 m. This seems like a pretty signifianct difference in whole plant height, but we can examine this difference further using a box-plot.

```{r}
# making boxplots
boxplot(americana_height)
boxplot(borbonia_height)
```

Let's make a t-test to further analyze the average whole plant height of the two species.

```{r}
# computing t-test
t.test(americana_height, borbonia_height)
```

Now that we've done our t-test, we can see that the p-value is very, very small, which means that the difference in whole plant height between Persea americana and Persea borbonia is very significant.
